---
title: "Project1"
team: "Andrew, Nicole, Ranjan"
---

```{r}

library(naniar)
library(caret)
library(ggplot2)
library(mlbench)
library(glmnet)
library(olsrr)
library(fmsb)
library(corrplot)
library(RColorBrewer)
library(funModeling)
library(tidyverse)
library(tidyr)
library(Hmisc)
library(kableExtra)
library(GGally)
library(plyr)
library(stringr)
library(agricolae) ## for cv
library(ggplot2)
library(GGally)
library(lindia)
install.packages("AICcmodavg")
library(AICcmodavg)
install.packages("Metrics")
library(Metrics)
```

```{r}

# Read in data
cars <- read.csv("data1.csv",stringsAsFactors = T)

# view data summary
summary(cars)

# view records with N/A's for review
cars_NA <- cars %>% filter_all(any_vars(is.na(.)))
cars_NA

# N/A's identified in Engine HP(69), Engine Cylinders(30), & Number of Doors(6)

```

```{r}

# Manually look up and add missing Engine HP values

cars_NA_EngineHP <- cars[is.na(cars$Engine.HP),]
cars_NA_EngineHP

# records 540 to 542 -> 2015, 2016, 2017 FIAT 500e -> 111 Hp
cars[c(540:542),5] <- 111

# records 2906 to 2909 -> 2017 Lincoln Continental (4 different trims) -> 305, 400, 335, 305 Hp
cars[c(2906:2909),5] <- c(305,400,335,305)

# records 4204 to 4207 -> 2017 Ford Escape (4 different trims) -> 168 Hp
cars[c(4204:4207),5] <- 168

# records 4706 to 4707 -> 2013, 2014 Honda Fit -> 123 Hp
cars[c(4706:4707),5] <- 123

# records 4915 to 4920 -> 2005 Ford Freestar (6 different trims) -> 193 Hp
cars[c(4915:4920),5] <- 193

# records 4786 to 4799 -> 2015, 2016, 2017 Ford Focus -> 143 Hp
cars[c(4786:4799),5] <- 143

# record 5779 -> 2014 Mitsubishi i-MiEV -> 66 Hp
cars[c(5779),5] <- 66

# records 5826 to 5841 -> 2015, 2016, 2017 Chevy Imapala (2 trims per year) -> 301 Hp
cars[c(5826:5841),5] <- 301

# records 6386 to 6395 -> 2014 to 2016 Nissan Leaf (3 trims per year) -> 107 Hp
cars[c(6386:6395),5] <- 107

# record 6579 -> 2015 Mercedes M-Class -> 200 Hp
cars[c(6579),5] <- 200

# record 6909 -> 2017 Lincoln MKZ (trim 1) -> 240 Hp
cars[c(6909),5] <- 240

# record 6911 -> 2017 Lincoln MKZ (trim 2) -> 350 Hp
cars[c(6911),5] <- 350

# record 6917 -> 2017 Lincoln MKZ (trim 3) -> 350 Hp
cars[c(6917),5] <- 350

# record 6919 -> 2017 Lincoln MKZ (trim 4) -> 400 Hp
cars[c(6919),5] <- 400

# records 6922 to 6925 -> 2014 Tesla Model S (4 trims) -> 362, 302, 691, 416 Hp
cars[c(6922:6925),5] <- c(362,302,691,416)

# records 6926 to 6930 -> 2015 Tesla Model S (5 trims) -> 323, 380, 691, 422, 380 Hp
cars[c(6926:6930),5] <- c(323,380,691,422,380)

# records 6931 to 6939 -> 2016 Tesla Model S (9 trims) -> 382, 315, 762, 422, 328, 422, 518, 691, 382 Hp
cars[c(6931:6939),5] <- c(382,315,762,422,328,422,518,691,382)

# record 8375 to 8376 -> 2013, 2014 Toyota RAV4 EV -> 154 Hp
cars[c(8375:8376),5] <- 154

# record 9851 to 9852 -> 2015 Kia Soul EV (2 trims) -> 154 Hp
cars[c(9851:9852),5] <- 154

# record 9853 to 9855 -> 2016 Kia Soul EV (3 trims) -> 109 Hp
cars[c(9853:9855),5] <- 109

# Verify no more NA's
cars_NA_EngineHP <- cars[is.na(cars$Engine.HP),]
cars_NA_EngineHP



# Manually look up and add missing "Engine Fuel Type" values
cars_blank <- cars %>% filter(Engine.Fuel.Type < 1)

# records 11323:11325 -> 2004 Suzuki Verona (3 trims)
cars[c(11322:11324),4] <- "regular unleaded"



```

```{r}

# Manually look up and add missing and "odd" Engine HP values
cars_NA_EngineCylinder <- cars[is.na(cars$Engine.Cylinders),]
cars_NA_EngineCylinder

# Input value of "Elec" for all electric vehicles (Engine.Fuel.Type = electric)
cars <- within(cars, Engine.Cylinders[Engine.Fuel.Type == 'electric'] <- 'Elec')

# Change Cylinder to categorical variable
cars$Engine.Cylinders <- as.factor(cars$Engine.Cylinders)

# Change Engine Cylinders value of Mazda RX-7 cars (rotary engine) to piston equivalent -> 6
cars[c(8696:8698),6] <- 6

# Change Engine Cylinders value of Mazda RX-8 cars (rotary engine) to piston equivalent -> 8
cars[c(8699:8715),6] <- 8

# Verify no more NA's
cars_NA_EngineCylinders <- cars[is.na(cars$Engine.Cylinders),]
cars_NA_EngineCylinders

```

```{r}

# Manually look up and add missing "Number of Doors" values
cars_NA_Doors <- cars[is.na(cars$Number.of.Doors),]
cars_NA_Doors

# record 4667 -> 2013 Ferrari FF has 2 Doors
cars[c(4667),9] <- 2

# records 6931 to 6935 -> 2016 Tesla Model S has 4 Doors
cars[c(6931:6935),9] <- 4

# Verify no more NA's
cars_NA_Doors <- cars[is.na(cars$Number.of.Doors),]
cars_NA_Doors

```

```{r}

# Verify ALL NA's have been changed
cars_NA <- cars %>% filter_all(any_vars(is.na(.)))
cars_NA
summary(cars)

```

```{r}

# Convert applicable categorical variables to "factors"
cars$Make <- as.factor(cars$Make)
cars$Engine.Fuel.Type <- as.numeric(as.factor(cars$Engine.Fuel.Type))
cars$Engine.Cylinders <- as.numeric(as.factor(cars$Engine.Cylinders))
cars$Transmission.Type <- as.numeric(as.factor(cars$Transmission.Type))
cars$Driven_Wheels <- as.numeric(as.factor(cars$Driven_Wheels))
cars$Vehicle.Size <- as.numeric(as.factor(cars$Vehicle.Size))
cars$Vehicle.Style <-as.numeric(as.factor(cars$Vehicle.Style))
cars$Number.of.Doors <-as.numeric(as.factor(cars$Number.of.Doors))
cars$Market.Category <-as.numeric(as.factor(cars$Market.Category))


# Convert applicable numerical variables to "numeric"
cars$Year <- as.numeric(cars$Year)
cars$Engine.HP <- as.numeric(cars$Engine.HP)
cars$highway.MPG <- as.numeric(cars$highway.MPG)
cars$city.mpg <- as.numeric(cars$city.mpg)
cars$Popularity <- as.numeric(cars$Popularity)
cars$MSRP <- as.numeric(cars$MSRP)

# Verify variable data types
str(cars)

# Too many levels to convert "Model" to Categorical/Factor.  This variable will not be considered in the analysis due to it's inherent correlation to "Make".

```

```{r}
# Review Market.Category values
levels(cars$Market.Category)
plyr::count(cars,'Market.Category')
# There are too many levels (72) to convert "Market.Category" to Categorical/Factor. The values of the variable are potential correlated to MSRP and Popularity which is the main consideration of this analysis so the variable will NOT be dropped.  Instead, the Market.Category variable will be read for the 11 unique string descriptors which make up all 72 combinations (levels).  11 new boolean/logical variables will be created and hold the value of whether or not the descriptor of each records shows up in its original "Market.Category" value.

pattern.Crossover <- paste("Crossover", collapse = "|")
pattern.Diesel <- paste("Diesel", collapse = "|")
pattern.Exotic <- paste("Exotic", collapse = "|")
pattern.Luxury <- paste("Luxury", collapse = "|")
pattern.Performance <- paste("Performance", collapse = "|")
pattern.FactoryTuner <- paste("Factory Tuner", collapse = "|")
pattern.FlexFuel <- paste("Flex Fuel", collapse = "|")
pattern.Hatchback <- paste("Hatchback", collapse = "|")
pattern.Hybrid <- paste("Hybrid", collapse = "|")

cars$Crossover <- grepl(pattern.Crossover, cars$Market.Category)
cars$Diesel <- grepl(pattern.Diesel, cars$Market.Category)
cars$Exotic <- grepl(pattern.Exotic, cars$Market.Category)
cars$Luxury <- grepl(pattern.Luxury, cars$Market.Category)
cars$Performance <- grepl(pattern.Performance, cars$Market.Category)
cars$FactoryTuner <- grepl(pattern.FactoryTuner, cars$Market.Category)
cars$FlexFuel <- grepl(pattern.FlexFuel, cars$Market.Category)
cars$Hatchback <- grepl(pattern.Hatchback, cars$Market.Category)
cars$Hybrid <- grepl(pattern.Hybrid, cars$Market.Category)

summary(cars)

```

```{r}

# visualize each variable against response variable (MSRP)
cars %>%
  gather(-MSRP, key = "var", value = "value") %>%
  ggplot(aes(x = value, y = MSRP)) +
    facet_wrap(~var, scales = "free") +
    geom_point()



# visualize change in car prices (MSRP) over time
cars %>%
  ggplot(aes(x = Year, y = MSRP)) +
    geom_point()

# There is a clear change (possibly in how the data is collected or calculated) in MSRP from years <= 2000 and >2000.  There are plenty of records on each side to complete our analysis so we will exclude all cars with "Year"<2000 as the "customer" is undoubtedly more interested in current/recent trends and correlations over the past 20 years rather than those older than 20 years.

cars2000 <- cars %>% filter(cars$Year>2000)

# 10,257 of the original 11,914 remain
summary(cars2000)

# There are no longer any records with Transmission.Type = "Unknown" so we can drop that level by running the as.factor again
cars2000$Transmission.Type <- droplevels(cars2000$Transmission.Type)


```
##Ranjan 
```{r}
#visualizing MSRP distrubution
hist(cars2000$MSRP,main="MSRP Distribution")  ### since this is heavily right skewed

## to get appropriate model I would like to remove high priced vehicle that is more than million
```

```{r}
cars2000 <- cars2000[cars2000$MSRP>9000,]
cars2000<- cars2000[cars2000$MSRP<1000000,]
dim(cars2000)

hist(cars2000$MSRP,main="MSRP Distribution")  ## this is better distribution of MSRP

```
   ## scatter plot for numeric variable
```{r}
pairs(~MSRP+Year+Popularity+Engine.HP+highway.MPG+city.mpg, data=cars2000)

##Engine HP and MSRP is correlated 
##city and highway mpg is highly correlated.

```
###dont include
```{r}
plyr::count(cars2000, 'Make')  
## only 3 genesis,2 plymouth,3 spyker, 5 mclaren ,5 alfa romeo . this might create problem when we split data.
par(mar = c(5.5,15,4.1,2.1))
barplot(plyr::count(cars2000,'Make')[,2], names.arg=plyr::count(cars2000,'Make')[,1], horiz=TRUE, col='red',las=1, main = "Make")


```

####    ##visualizing categorical variable#########################
```{r}
par(mar = c(5.5,15,4.1,2.1))
barplot(plyr::count(cars2000,'Engine.Fuel.Type')[,2], names.arg=plyr::count(cars2000,'Engine.Fuel.Type')[,1], horiz=TRUE, col='red',las=1, main = "Engine fuel type")

## more number of cars with regular unleaded gas
```

```{r}
par(mar = c(5.1,4.5,4.1,2.1))
barplot(plyr::count(cars2000,'Engine.Cylinders')[,2], names.arg=plyr::count(cars2000,'Engine.Cylinders')[,1], horiz=TRUE, col='red',las=2, main = 'Number of Cylinders')

### 4,6,8 are highest cyclinders
```


```{r}
par(mfrow = c(2,2))
par(mar = c(5.1,15,4.1,2.1))
barplot(plyr::count(cars2000,'Transmission.Type')[,2], names.arg=plyr::count(cars2000,'Transmission.Type')[,1], horiz=TRUE, col='red',las=2, main = 'Transmission Type')
barplot(plyr::count(cars2000,'Number.of.Doors')[,2], names.arg=plyr::count(cars2000,'Number.of.Doors')[,1], horiz=TRUE, col='red',las=2, main = 'Number of Doors')
barplot(plyr::count(cars2000,'Driven_Wheels')[,2], names.arg=plyr::count(cars2000,'Driven_Wheels')[,1], horiz=TRUE, col='red',las=2, main = 'Driven Wheels')
barplot(plyr::count(cars2000,'Vehicle.Size')[,2], names.arg=plyr::count(cars2000,'Vehicle.Size')[,1], horiz=TRUE, col='red',las=2, main = 'Vehicle Size')

```


```{r}
par(mar = c(5.1,10,4.1,2.1))
barplot(plyr::count(cars2000,'Vehicle.Style ')[,2], names.arg=plyr::count(cars2000,'Vehicle.Style ')[,1], horiz=TRUE, col='red',las=2, main = 'Vehicle Style ')

```
#######
```{r}
# visualize change in car prices (MSRP) since 2000

cars2000 %>%
  ggplot(aes(x = Year, y = MSRP)) +
    geom_point()

# take log(MRSP)
cars2000 %>%
  ggplot(aes(x = Year, y = log(MSRP))) +
  geom_point()

# visualize all individual variable relationships to MSRP

# engine.hp
cars2000 %>%
  ggplot(aes(x = Engine.HP, y = log(MSRP))) +
  geom_point()

##highway.mpg ###log looks better
cars2000 %>%
  ggplot(aes(x =log(highway.MPG), y = log(MSRP))) +
  geom_point()

##audi A6 row  837 from 354mpg  to 34 ##error 
cars2000[c(837),13] <- 34
view(cars2000)

cars2000 %>%
  ggplot(aes(x =highway.MPG, y = log(MSRP))) +
  geom_point()

#citympg

cars2000 %>%
  ggplot(aes(x = city.mpg, y = log(MSRP))) +
  geom_point()

##popularity

cars2000 %>%
  ggplot(aes(x =  Popularity, y = log(MSRP))) +
  geom_point()

##nicole code
pairs((cars2000[,c(3,5,6,13,14,15,16)]))
ggpairs((cars2000[,c(3,5,6,13,14,15,16)]))
plot_num(cars2000)




```

```{r}
# set seed
set.seed(123)

# Split data into smaller sets: 80% training, 10% test, and 10% validate
ss <- sample(1:3,size=nrow(cars),replace=TRUE,prob=c(0.8,0.1,0.1))
train <- cars2000[ss==1,]
test <- cars2000[ss==2,]
valid <- cars2000[ss==3,]

#summary(train)
#summary(test)
#summary(valid)

```
## full model 
```{r}
Fullmodel = lm (log(MSRP)~.-Make -Model,data = train)
summary(Fullmodel)
par(mfrow = c(2,2))
plot(Fullmodel)
CV(Fullmodel)


```

## CV           AIC          AICc           BIC         AdjR2 
 
        ##  Inf -4.024610e+04 -4.005664e+04 -3.436392e+04  9.808313e-01 


```{r}
# Fullmodelselect = ols_step_both_aic(Fullmodel,details=TRUE)
# summary(Fullmodelselect)
```

###linear regression for numeric variables
```{r}
linearN <- lm(log(MSRP)~Year+Popularity+Engine.HP+highway.MPG+city.mpg+, data=train)
summary(linearN)
vif(linearN)
summary(linearN)$r.squared

## there city mpg and highway mpg is highly correlated .so either one of the variable can removed to make prediction
```

```{r}
par(mfrow = c(2,2))
plot(linearN)


```


##without city mpg
```{r}
linearN1 <- lm(log(MSRP)~Year+Popularity+Engine.HP+city.mpg+Engine.Fuel.Type+Driven_Wheels+Engine.Cylinders+Make+highway.MPG+Transmission.Type+Vehicle.Size+Vehicle.Style+, data=train)
summary(linearN1)

# summary(linearN1)$r.squared
# par(mfrow = c(2,2))
# plot(linearN1)
gg_diagnose(linearN1)
```


###selection of variable
```{r}
linearn1select =ols_step_both_p(Fullmodel, prem = 0.01, pent = 0.02, details = F, progress = T)
summary(linearn1select)

```

```{r}
linearn1F =ols_step_forward_p(Fullmodel, penter = 0.02, details = TRUE)
summary(linearn1F)

```

```{r}
linearnB = ols_step_backward_p(Fullmodel,fit, prem = 0.01,details=TRUE)
summary(linearnB)

```

```{r}
#predictforward <- predict(linearn1F,newdata=test)

```

```{r}
modelstep <- lm(log(MSRP)~Performance + Transmission.Type + Year + Diesel 
+ Vehicle.Style 
+ Hatchback 
+ Hybrid 
+ Driven_Wheels 
+ FactoryTuner 
+ city.mpg 
+ Engine.Cylinders 
+ Crossover 
+ Engine.Fuel.Type 
+ FlexFuel 
+ Market.Category 
+ Popularity 
+ highway.MPG,data=train)
summary(modelstep)
```

```{r}
modelforward <-lm(log(MSRP)~ Engine.HP 
+ Exotic 
+ Luxury 
+ Performance 
+ Transmission.Type 
+ Year 
+ Diesel 
+ Vehicle.Style 
+ Hatchback 
+ Hybrid 
+ Driven_Wheels 
+ FactoryTuner 
+ city.mpg 
+ Engine.Cylinders 
+ Crossover 
+ Engine.Fuel.Type 
+ FlexFuel 
+ Market.Category 
+ Popularity 
+ highway.MPG,data=train)
summary(modelforward)
```
##backward removed Number of doors and vehicle size
```{r}
modelbackward <- lm(log(MSRP)~.-Number.of.Doors-Vehicle.Size-Model-Make,data=train) 
summary(modelbackward)
```

```{r}
#define list of models 
models <- list(modelstep, modelforward, modelbackward)

#specify model names
mod.names <- c('stepwise_selection', 'forward_selection', 'backward_selection')

#calculate AIC of each model
aictab(cand.set = models, modnames = mod.names)

```

```{r}
Predict_Stepwise = predict(modelstep,test)
Predict_Stepwise = exp(Predict_Stepwise)
Predict_Stepwise = as.data.frame(Predict_Stepwise)
StepwisePredicted = cbind(test$MSRP,Predict_Stepwise)
names(Predict_Stepwise)[1] = "test"
names(Predict_Stepwise)[2] = "prediction"


```

```{r}
confusionMatrix(Predict_Stepwise$prediction,reference = Predict_Stepwise$test)

str(Predict_Stepwise)
str(test$MSRP)

sse(test$MSRP,Predict_Stepwise$prediction)




